---
title: "Project Cleaning"
author: "Sanjana Godiawala"
date: "23/11/2021"
output:
  pdf_document: default
  html_document: default
---
## Dataset
```{r}
library(dplyr)
library(missForest)
library(tidyr)
```

```{r}
ds <- read.csv("dataset_train.csv")
head(ds)
```
## Dimension of the dataset
```{r}
dim(ds)
```

## Structure of the dataset

```{r}
str(ds)
```
## Dataset Summary 
```{r}
summary(ds)
```
## Checking for nul values
```{r}
sum(is.na(ds))
```
## Since there are only 6 nul values omiting these rows. 
```{r}
ds <- na.omit(ds)
sum(is.na(ds))
```
## Printing the head and tail of the dataset
```{r}
head(ds)
tail(ds)
#Random 10 records
sample_n(ds,10)
```
## Factoring the columns vendor id, pick loc, drop loc,number of passenger,payment method, rate and store flag for categorising the data values as it is more favorable for processing. 

```{r}
ds$vendor_id <- factor(ds$vendor_id)
ds$pickup_loc <- factor(ds$pickup_loc)
ds$drop_loc <- factor(ds$drop_loc)
ds$num_passengers <- factor(ds$num_passengers)
ds$payment_method <- factor(ds$payment_method)
ds$rate_code <- factor(ds$rate_code)
ds$stored_flag <- factor(ds$stored_flag)
str(ds)
```
## Change the type of driver tip and total_amount to numeric

```{r}
ds$driver_tip <- as.numeric(ds$driver_tip)
ds$total_amount <- as.numeric(ds$total_amount)
str(ds)
summary(ds)
```

## Numeric conversion 
```{r}
ds$mta_tax <- as.numeric(ds$mta_tax)
ds$extra_charges <- as.numeric(ds$extra_charges)
ds$improvement_charge <- as.numeric(ds$improvement_charge)
ds$toll_amount <- as.numeric(ds$toll_amount)
```

## Date conversion
```{r}
ds <- ds %>% separate(pickup_time,c("Pickup_date","Pickup_Time"),sep=" ")
ds <- ds %>% separate(drop_time,c("Drop_date","Drop_Time"),sep=" ")
ds$Pickup_date <- as.Date(ds$Pickup_date,format = "%d-%m-%Y")
ds$Drop_date <- as.Date(ds$Drop_date,format = "%d-%m-%Y")
ds$Drop_Time <-strptime(ds$Drop_Time, format = "%H:%M")
ds$Pickup_Time <- strptime(ds$Pickup_Time,format="%H:%M")
str(ds)
```

```{r}
ds <- na.omit(ds)
sum(is.na(ds))
summary(ds)
## removing duplicate rows
ds <- distinct(ds)
dim(ds)
```

## Adding a column Duration which is difference of pick-up and drop-off time
```{r}
ds$Duration <- difftime(ds$Drop_Time, ds$Pickup_Time, units = "mins")
head(ds)
```

## storing the cleaned dataset

```{r}
write.csv(ds,"cleaned_dataset.csv",row.names = FALSE)
```
